---
title: "BUS212 6b Automated Data Collection"
output:
  word_document: default
  html_notebook: default
---
## Introduction

This code is adapted from the book *AUTOMATED DATA COLLECTION WITH R*, by Simon Munzert, Christian Rubba, Peter Meissner, Dominic Nyhuis. The example comes from Chater 1, and demonstrates how to read in data from a website, prepare it for analysis, and then produce some visualization.

The example uses the Wikipedia page about  the _Endangered World Heritage Sites_, which are places in the world that are (a) included in the list of antiquities and natural wonders known as World Heritage Sites and (b) in danger due to natural or political reasons.

The script requirs four packages:

```{r, warning=FALSE, error=FALSE}

# load packages
library(stringr)
library(XML)
library(maps)
library(RCurl)
```

The relevant website is https://en.wikipedia.org/wiki/List_of_World_Heritage_in_Danger. The first code chunk reads the relevant data into an object called `heritage_parsed` using the command `getURL`.
```{r}
# parsing from Wikipedia web site
fileURL <- "https://en.wikipedia.org/wiki/List_of_World_Heritage_in_Danger"
xData <- getURL(fileURL)
heritage_parsed <- htmlParse(xData)
```

The heritage_parsed object contains the full list of sites as well as those in danger. The next few lines separate the full table and the table of endangered sites.

```{r}
tables <- readHTMLTable(heritage_parsed, stringsAsFactors = FALSE)
danger_table <- readHTMLTable(heritage_parsed, stringsAsFactors = FALSE, which = 2) # alternatively: directly select second table

# extract desired table
danger_table <- tables[[2]]
names(danger_table)

# select and rename columns
danger_table <- danger_table[,c(1,3,4,6,7)]
colnames(danger_table) <- c("name","locn","crit","yins","yend")
danger_table$name[1:3]
```
In its raw form, the data are not ready for analysis. This chunk performs several steps to clear and reorganize the data.  Towards the end of the chunk, notice that it is pulling out longitude and latitudes in order to map the sites.  

```{r}
# cleanse criteria
danger_table$crit <- ifelse(str_detect(danger_table$crit, "Natural")==T, "nat", "cult")

# cleanse years
danger_table$yins <- as.numeric(danger_table$yins)
danger_table$yins
danger_table$yend

yend_clean <- unlist(str_extract_all(danger_table$yend, "^[[:digit:]]{4}"))
danger_table$yend <- as.numeric(yend_clean)
danger_table$locn[c(1,3,5)]

# get countries
reg <- "[[:alpha:] ]+(?=[[:digit:]])"
country <- str_extract(danger_table$locn, perl(reg)) # use forward assertion in Perl regular expression
head(country)
country[29] <- "Côte d'Ivoire / Guinea"
country[32] <- ""
danger_table$country <- country

# get coordinates
reg_y <- "[/][ -]*[[:digit:]]*[.]*[[:digit:]]*[;]"
reg_x <- "[;][ -]*[[:digit:]]*[.]*[[:digit:]]*"
y_coords <- str_extract(danger_table$locn, reg_y)
(y_coords <- as.numeric(str_sub(y_coords, 3, -2)))
danger_table$y_coords <- y_coords
x_coords <- str_extract(danger_table$locn, reg_x)
(x_coords <- as.numeric(str_sub(x_coords, 3, -1)))
danger_table$x_coords <- x_coords
danger_table$locn <- NULL
```
With data preparation completed, here is a map of the  locations of the endangered sites.
```{r}
par(oma=c(0,0,0,0))
par(mar=c(0,0,0,0))
pch <- ifelse(danger_table$crit == "nat", 19, 2)
map("world", col = "darkgrey", lwd = .5, mar = c(0.1,0.1,0.1,0.1))
points(danger_table$x_coords, danger_table$y_coords, pch = pch, col = "black", cex = .8)
box()
```
Next, the code creates a histogram showing the years in which sites became endangered.

```{r}

# table heritage criteria
table(danger_table$crit)

# plot year of endangerment

par(oma=c(0,0,0,0))
par(mar=c(4,4,1,.5))
hist(danger_table$yend, freq=TRUE, xlab="Year when site was put on the list of endangered sites", main="")
box()
```

And finally, a graph showing time between inscription and endangerment.

```{r}
duration <- danger_table$yend - danger_table$yins
par(oma=c(0,0,0,0))
par(mar=c(4,4,1,.5))
hist(duration, freq=TRUE, xlab="Years it took to become an endangered site", main="")
box()
```






```

